---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ProgressDataDirectConnect4R

<!-- badges: start -->

[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental) [![CRAN status](https://www.r-pkg.org/badges/version/ProgressDataDirectConnect4R)](https://CRAN.R-project.org/package=ProgressDataDirectConnect4R)

<!-- badges: end -->

Package has three main goals.

-   Provide an OpenEdge DBI driver for R.

-   Support `dbplyr` generics.

-   Allow for future expansion using other DataDirect Connect Java connectors.

## Installation

You can install the released version of ProgressDataDirectConnect4R from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("ProgressDataDirectConnect4R")
```

Install the development branch using

``` r
remotes::install_github("data-science-r/ProgressDataDirectConnect4R@develop")
```

Once loaded, you can load the OpenEdge JDBC driver using

``` r
ProgressDataDirectConnect4R::OpenEdge()
```

or alternatively

``` r
DBI::dbDriver("OpenEdge")
```

for indirectly finding the driver using `DBI`. Both return a `JDBCDriver`, an S4 wrapper for the underlying Java driver that can serve as a driver argument for DBI connections.

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(ProgressDataDirectConnect4R)
## basic example code
```

## Requirements

Needs Java and for Windows you also need `JAVA_HOME`. On Windows you can use Chocolatey to install AdoptOpenJDK 11, or some later version. You might see messages such as the error below.

    error: JAVA_HOME cannot be determined from the Registry

You can set up the environment variable directly from R on Windows using

``` r
Sys.setenv(JAVA_HOME = "C:/Program Files/AdoptOpenJDK/jdk-11.0.11.9-hotspot")
```

## Dependencies

Depends on `RJDBC` which itself depends on `DBI` and `rJava` packages. It appears in the package description under `Depends` rather than `Imports`. Loading this package loads `RJDBC` as a pre-requisite.

## Downloads

Downloads the evaluation [Jar](https://media.datadirect.com/download/files/evals/JDBC/PROGRESS_DATADIRECT_JDBC_OE_ALL.jar) from Progress. Temporarily unpacks the resources.

The underlying Java JAR comes from Progress.

Progress supply the drivers under 15-day evaluation. However, the JAR continues to work after 15 days without trouble. It was originally downloaded in June 2021 as the DataDirect Connect for JDBC 5.1 SP4. So far, so good.

## Continuous Integration

Uses `rocker/ml-verse` for testing and coverage. This image incorporates the `rJava` package and the Java Run-time Environment. The image has a disadvantage: the size is big, around six and a half gigabytes. Pulling it slows down integration steps.

Note that package checking requires no errors *and* no warnings for success. Job success allows for only "notes."
